A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-09-27 reviewer_response.tex

\documentclass{article} %{aiaa-tc}

	\addtolength{\oddsidemargin}{-.875in}
	\addtolength{\evensidemargin}{-.875in}
	\addtolength{\textwidth}{1.75in}

	\addtolength{\topmargin}{-.875in}
	\addtolength{\textheight}{1.75in}

\date{}
\setcounter{secnumdepth}{0}

\title{\large Authors' Responses to the Reviewers for the Manuscript,\\
\large ``A Graph Theoretic Approach to Problem Formulation for
Multidisciplinary Design Analysis and Optimization'' \\  \vspace{0.25in} \normalsize (Reviewer comments are set in red text and the authors' responses are set in black text)}

@@

\usepackage{color}

%\newenvironment{rev}{\vspace{2em}\itshape \textcolor{red}}{}
\newenvironment{rev}{\vspace{2em} \textcolor{red}}{}

@@

\rev{This idea of PSG was mentioned but was not covered in
the paper. For completeness and value proposition sake, I would recommend to
add at least half section to show how it will help and look like. You can
take the simplest formulation of your choice and show how MCG and FPG will
translate and look as a PSG. A brief discussion of advantages vs
disadvantages will be useful. }

Although we feel that discussion of the PSG in significant depth would be too lengthy for the present paper, we agree that the topic does warrant a more complete discussion than that provided in the
original draft.  We have added subsections 3.5 and 4.4 to address the the driver nodes and driven edges which
together distinguish a PSG from an FPG.  We have also added subsection 5.5,
which provides an example of a simple PSG and compares and contrasts it with an XDSM diagram.

\rev{ The idea of “holes” and “collision” is explained a few times in the paper.
It can be reduced. }

The authors agree that the concepts of holes and collisions were discussed too often.  The terms are formally defined in section 4.1. Section 5.3 provides a mathematical description of the set of nodes that meet
the criteria for holes and collisions in the context of the FPG algorithm, and it is our opinion that this discussion should stand separately from the definitions.

We have checked sections 4.1 and 5.3 and have removed most instances of redundancy in other parts of the paper, including informal definitions repeated several times in the text.  Nonetheless, we feel that preserving the separate mathematical discussions in sections 4.1 and 5.3 makes for a clearer message in the paper. The concept is somewhat subtle since, in our definitions, a node can be either a design variable or a hole, but not both. Similarly, a collision occurs when a variable has two incoming edges if the problem is single fidelity but does not occur in this situation if the problem is multi-fidelity. The additional discussion is intended to clarify theses subtleties.

@@

\rev{The review of other graph based techniques is good and highlights their
limitations, but I would like to see a discussion of what challenges a
practitioner may face while trying to use the approach suggested in this
paper. The MCG and FPG will cover lot of information like objectives and
constraints, but at what expense. Comparing XDSM with MCG or FPG, one might
find XDSM as more easy to read and comprehend. For a similar size problem,
FPG looks far more complex and bigger.}

We agree that XDSM is more visually informative and concise. However, an easily human-readable format was not the
goal of the new syntax. Instead, the goal is to create a graph structure that
can be manipulated algorithmically to achieve feasible MDO data flows. As a consequence of this goal, the graph grows larger and becomes less human readable. We believe that this tradeoff is acceptable for two reasons:

\begin{enumerate}
    \item It should be possible to convert between the gray lines of the XDSM format and the proposed syntax
    in a lossless manner.
    It is our conclusion that the PSG in our syntax represents a subset of the information in the XDSM diagram; namely, the data connectivity elements of XDSM and PSG are equivalent graphs.  The implication is that it would be possible to write an algorithm and associated code to ``abstract'' the FPG into a partial XDSM diagram that could then be elaborated by the user to add the solution-specific elements.  It may then be possible to write a PSG from an XDSM, if the additional information about all of the individual variable connections from the FPG graph were preserved in the background.
    (This is true for the gray lines of XDSM but not the black ones. Only data connections are described in our new syntax.)
    \item Algorithmic methods are a useful way for analyzing MDO problem structure, so it seems appropriate to have a graph structure that works well for this use, even at the expense
    of human readability.
\end{enumerate}

Please see the revised end of the introduction (Section 1) in which the algorithmic purpose of the new graph syntax is now stated more clearly.  We have also added the following text in the last paragraph of Section 2.2 to clarify the issue:

\begin{quotation} 
\textit{REMS and XDSM provide excellent
human readable formats for an MDAO problem description, whereas the syntax described in this paper is designed to provide an effective machine readable format
for MDAO problem definition. In the interest of simplifying algorithmic operations on the graphs, the specified format is both more verbose and less visually informative. The benefit is a consistent and easily utilized
graph from a algorithmic perspective.}
\end{quotation}

@@

We have added  Fig. 8 and Fig. 13 and associated discussion to address this request.

\rev{Is it really adding a lot of value to have graphs like MCG or FPG which
are very explicit in nature and can become difficult to read}

As discussed above, we believe that the value of the MCG and FPG is not derived from their ability to represent the data in a human readable
form, but in their straightforward translation to machine readable formats. Section 5.3 introduces an
algorithm for testing whether MDO data connectivity graphs are FPGs and for reducing graphs to FPGs. As problems grow larger and more complex,
this type of algorithm is expected to become ever more valuable.

Such an algorithm could be implemented into a framework or stand-alone tool to help teams to develop the
specifics of a design problem before they begin detailed development work.
Section 6 shows that for even simple problems, the graph is
non-trivial. As problems grow even a small amount from the size of the example, an algorithmic approach for studying problem structure becomes invaluable. In sum, it is our belief that the graph syntax derives its usefulness from
its effectiveness in describing MDO problem graphs in computational syntax.

\rev{Can the graph be compressed in a logical way to reduce the number of nodes
and their interconnections while still conveying important and critical
information for that stage (MCG stage or FPG stage)? Reducing cycles is
discussed but the graph may still look complex.
\\ \indent
Looking at these graphs (for example Fig. 8), a node for example Z2
appears several times. It may be useful to figure out holes and collision,
but once it is resolved, can those nodes be consolidated to provide simpler
visualization ?}

@@

The answer to this question is not trivial, as it depends on a number of factors such as the number of analyses and the number of variables in those analyses. However, roughly speaking, one could measure the complexity of a graph by the total number of nodes and edges. The number of nodes grows roughly linearly with the number of analysis tools and the number of variables. The number of fixed edges grows linearly with the number of variables as well. The number of connection edges depends entirely on the specifics of the problem, but, since any variable will have only a small finite number of incoming edges (usually 1, but maybe 2 or 3 in multi-fidelity case) one can avoid an unbounded growth. Therefore, very roughly speaking, the graph will grow as O(N+M) where N is the number of variables and M is the number of analyses.

If one assumes that large problems have hundreds of components and millions of variables, then graphs with a few million nodes and edges are possible. These numbers are very large, but are not outside of the realm of what is readily manageable and rapidly evaluated by today's graph libraries in common programming languages.

\rev{
I will assume that creation of such a graph and interconnections will be
expected to be done by some kind of software tool or wizard. A brief
discussion about why this MCG, FPG and PSG idea is better to lead to such a
system can be useful. Although this itself can be a separate research and
paper, but why this idea can provide a better foundation can be valuable.
}

Please see the new section 2.4 for a brief discussion of our envisioned usage
of the graph language in a software tool. We have also mentioned in Section 5.3 that the algorithm can be straightforwardly implemented in any programming language capable of managing lists, such as Python.

@@

\rev{Minor comments:
Section 1 Line 46: I think other commercial tools such as iSight,
modeFRONTIER should also be mentioned since their usage is more widespread
compared to OpenMDAO.}

Agreed. We have now mentioned these tools alongside OpenMDAO in  Section 1.

@@

\rev{Page 3, Line 21, 2nd column: change ``descrition'' to ``description''}

Fixed.

\rev{Page 4, Line 37, 1st column: what does ``REMS'' stand for?}

``Reconfigurability in MDO Problem Synthesis''. The full name has been added to the paper to define the acronym.

@@

\rev{Page 4, Line 45, 1st column: change ``REMS address'' to ``REMS addresses''}

Fixed.

\rev{Page 4, Lines 35 to 55, 1st column:  This paragraph is a bit awkward in
structure.  In the second to last sentence you state, "REMS provides syntax
for variables and functions but does not facilitate inclusion of solvers or
optimizers in the graph."  The reader then expects your method to address
these shortcomings in the next sentence but instead you state, "...the goal
of allowing graphs to describe analysis couplings and more closely interact
with existing design processes and MDAO frameworks," which you never really
called out as a shortcoming of REMS in the first place.  I think you could
improve the flow/wording here.}

The authors agree that this paragraph was poorly worded. We have re-worded it to attempt
to make the comparison with REMS more clear. The text now states (last two sentences, first column of p. 4): ``... REMS does not provide a mechanism for inclusion of solvers or optimizers in the graph. This fundamentally limits REMS from describing the specifics of different solution strategies as applied to a particular problem.''  We have removed the oddly juxtaposed statement about the goal of our graph syntax that was noted by the reviewer from the paragraph.  However, based on the comments by Reviewer 1 noted above, we have now added an additional description of how our method can capture optimizers and solvers in the context of the problem solution graph (PSG).

@@

The authors agree.  We have added equation numbers (now Eqns. 8 and 9). We have also provided verbal descriptions to accompany the set notations.

\rev{Lines 5 - 12, 2nd column, Page 12:  Just a general question spurred by the
text here.  What criteria would be applied to determine which FPG is better
than another?}

After considerable deliberation, we chose not to address the issue of how one should select one FPG in preference to another  in the context of the present paper. The answer
to this question depends heavily on the specific situation. If one is limited by run-time of the MDO workflow,
a smaller FPG with fewer cycles and variables may be preferable. If one is trying to resolve a detailed physical question, e.g. subtle design changes associated with nacelle-wing interference, then it may be preferable  to select a
more high fidelity and computationally expensive FPG. These notions are introduced in the example problem, but further development was beyond the scope of the paper.

The MCG provides a good foundation for identifying multiple possible FPGs. The user can then select the best FPG for the present stage of the design process based on his/her own criteria.  This decision can be revisited later in the process as the design goals and associated criteria change.

@@

\rev{Line 57, 1st column, Page 13: change ``as a input'' to ``as an input''}

Fixed.

\rev{Now for the request of more significant importance.  This is coming from
someone with significant experience in MDO and who uses a lot of the
analysis tools you specified in your case study on a daily basis.  I really
liked this article but felt it was lacking some punch in the results
section.  This was namely in the form of quantitative results.  You vaguely
give some in the form of the different resulting connectivity graphs and
this was good but to really drive home the value of your research to a
potential user I think you could do a bit better with very minimal work.
The potential user of this methodology will be concerned with one of two
things: 1) minimizing computational time or 2) increasing the accuracy of
the results.  I think you've given examples of these two scenarios already
in figures 15 and 16, so why not provide a table of the CPU times (1) and/or
deviation from higher order analyses (2) for each of these graphs.  Doesn't
have to necessarily be these two metrics but something to really tie the
output of your methodology to something tangible for comparison purposes.
You could even include the metrics for a randomly obtained FPG as a
baseline.  Then the person reading the article can say, "Wow, given a set of
analysis tools and a general problem, I can employ this method to obtain an
FPG that significantly reduces my overall computational time." or "Wow,
given a set of analysis tools and a general problem, I can employ this
method to obtain an FPG that significantly improves my overall computational
accuracy while keeping my computational efficiency as high as possible."}

We agree that this sort of quantitative comparison would be very valuable. However, we
believe that, in order not to be misleading, this comparison needs to be larger in scope than just that which could be obtained
by executing the example problem currently presented in the paper. We believe that one would need to address a wide range of problems with different graph structural
characteristics and different inherent nonlinear features in the analyses in order to make statements that are more generalizable. The challenge is that quantitative measures are inherently
dependent on particular details of a problem.
For example, for coupled analyses, total CPU time is highly dependent on the number of iterations required to converge the system of nonlinear functions to a consistent solution within a specified tolerance.  This convergence is tied very closely to the specifics of the coupling and the nonlinearity.  We had previously considered the idea of selecting ``optimal FPGs'' in terms of aggregate metrics that could be specified \textit{a priori} as meta-data for a particular analysis, but defining these metrics without considering problem coupling and nonlinearity proved especially challenging.
To address this issue well would therefore necessitate the evaluation of many example problems with different characteristics.  This work, while very valuable, would
likely require an entire paper in its own right. We thank the reviewer for this suggestion, and we consider it to be a great topic for future work in this area.

@@

\rev{If you are worried about length, I think the conclusions section may be cut
in half and still accomplish its intended purpose.}

We agree that the conclusion section
could be more concise. We have shortened it by two paragraphs and improved the wording.  In particular, we have removed the unnecessary recap of collisions and holes that was noted by Reviewer 1.

@@

The reviewer has correctly identified the major motivation for this work. The goal is indeed to
assist engineers and teams of engineers in the process of building a model out of a set of interconnected
tools.
We believe that the distinction between phases 1 and 2 in the reviewer's  comments are covered adequately
in the the introduction and especially in the notional problem described in the indented text on p. 2.
However, to clarify the distinction between phases 2 and 3 indicated by the reviewer, we modified the text following the indented text at the bottom of p. 2 to read as follows:

\begin{quotation}
    \textit{Our goal is to develop formalism for expressing analysis interconnectivity and for determining feasible
    analysis tool sets to assist an engineering team conducting this task. Because the problem deals with
    interconnectivity, we base our approach on the representations and techniques of graph theory.
    The approach begins by constructing the \emph{maximal connectivity graph (MCG)} describing all possible
    interconnections between the analysis tools proposed by the engineers. Graph operations are then
    conducted to reduce the MCG to a \emph{fundamental problem graph (FPG)} that describes a set of analysis
    tools that collectively specify a system-level design problem. The concept of the FPG and the identification of feasible FPGs from an MCG are the main contributions of the paper.}

   \textit{ The information in an FPG represents the engineering design problem that needs to be solved, but the FPG does not
    itself provide sufficient information to run a design optimization. In particular, it does not provide information
    about \textit{how} to solve the problem. Even after the problem is defined, an integration platform or framework needs
    to be selected and appropriate optimization methods identified. This last step essentially
    involves converting the FPG into a usable model. We briefly consider methods to further
    manipulate the FPG into a \emph{problem solution graph (PSG)} which could be useful in this
    task, but the PSG is not the primary focus of this paper. Rather, the work is primarily concerned with applying graph
    theory to the creation of an FPG and the benefits to the design process obtained by doing so.}
\end{quotation}

### 2013-09-27 conclusions.tex
Submitted version of revisions 1

\section*{Conclusions}
In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory. 
Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, 
DSM, and XDSM, but it provides new constructs tailored to algorithmic analysis and 
manipulation. The proposed syntax has applications in early phases of MDAO problem 
formulation, especially for problems with large numbers of variables and analyses. 
In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and 
the Fundamental Problem Graph (FPG).

The MCG addresses the question, ``Given a set of analysis tools, 
what are all of the variable interconnections between them that could be established?''  
The MCG provides a structured formalism to identify \textit{holes} and 
\textit{collisions} in interconnections between the set of analysis tools. 
In order to achieve a valid problem formulation, all of the holes and collisions must 
be resolved. We present an algorithm to perform this resolution which guarantees that an FPG will be found if one exists. In some cases, this algorithm relies on additional input from the 
designers to make decisions. One benefit of a graph theoretic approach is the standard 
algorithms that can be used to inform the user's decisions, such as cycle detection, 
minimum spanning trees, and shortest path algorithms.

@@

>>>>

\section*{Conclusions}
In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory.
Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT,
DSM, and XDSM, but it provides new constructs tailored to algorithmic analysis and
manipulation. The proposed syntax has applications in early phases of MDAO problem
formulation, especially for problems with large numbers of variables and analyses.

In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and
the Fundamental Problem Graph (FPG).
The MCG addresses the question, ``Given a set of analysis tools,
what are all of the variable interconnections between them that could be established?''
The MCG provides a structured formalism to identify \textit{holes} and
\textit{collisions} in interconnections between the set of analysis tools.
In order to achieve a valid problem formulation, all of the holes and collisions must
be resolved. We present an algorithm to perform this resolution which guarantees that an FPG will be found if one exists. In some cases, this algorithm relies on additional input from the
designers to make decisions. One benefit of a graph theoretic approach is the standard
algorithms that can be used to inform the user's decisions, such as cycle detection,
minimum spanning trees, and shortest path algorithms.

@@

%Techniques such as REMS and $\Psi$ focus on developing partitioned subproblems corresponding a given fundamental problem formulation.

Our graph-based approach is intended primarily to infer possible fundamental problem formulations, via their corresponding FPGs.  However, the syntax also facilitates the development of a Problem Solution Graph (PSG) that describes the MDAO partitioning and coordination strategy based on a specified FPG.  The formulation of a PSG could take advantage of other methods such as $\Psi$, REMS, and XDSM to assist in the specification of subproblem partitions and the addition of coordination-specific elements such as drivers and compatibility constraints.

### 2013-09-27 intro.tex
Submitted version of revisions 1

    Multidisciplinary Design Analysis and Optimization (MDAO)
    frameworks such as iSight\textsuperscript{\textregistered},\\ ModelCenter\textsuperscript{\textregistered}, ModeFrontier\textsuperscript{\textregistered}, and OpenMDAO\cite{Gray2012} have enabled a new level of analysis tool integration 
    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
    new capability for tool integration has created a new challenge. Even models with tens of analyses could include hundreds or thousands
    of variables that are interdependent and must be linked in the framework.

@@

    Lamb and Martins also included objectives and constraint functions as nodes 
    in an Extended DSM (XDSM) in order to capture a more complete description 
    of solution strategies for MDAO problems \cite{Lambe2012}. Unlike REMS, 
    XDSM also includes nodes for solvers and optimizers to enable complete 
    definition of MDAO architectures. Martins uses XDSM to describe thirteen different 
    optimization architectures in a survey paper that provides a novel 
    classification of the different techniques \cite{Lambe2011}. With the 
    additional information included in an XDSM, Lu and Martins applied both 
    ordering and partitioning algorithms to an MDAO test problem termed the 
    Scalable Problem \cite{Lu2012}.

>>>>

    Multidisciplinary Design Analysis and Optimization (MDAO)
    frameworks such as iSight\textsuperscript{\textregistered},\\ ModelCenter\textsuperscript{\textregistered}, ModeFrontier\textsuperscript{\textregistered}, and OpenMDAO\cite{Gray2012} have enabled a new level of analysis tool integration
    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This
    new capability for tool integration has created a new challenge: Even models with tens of analyses could include hundreds or thousands of variables that are interdependent and must be linked in the framework.

@@

Tosserams et al. developed a technique for problem partitioning named $\Psi$ that, like REMS, is based on a formal language specification \cite{Tosserams2010}.  The $\Psi$ method allows the user to specify how an MDO problem is to be decomposed into multiple optimization subproblems.  These subproblems may involve one or several analyses and multiple design variables.  Unlike REMS which builds subproblems automatically, $\Psi$ is intended to enable a manual process of subproblem construction based on the designers’ goals.  After subproblems are built, the language includes constructs to link them together to form into a system-level optimization problem statement.  $\Psi$ does not predispose the coordination strategy in the MDO architecture; any applicable approach can be used to achieve the required consistency among subproblems during the system level optimization.   Compiler and generator functions have been developed for $\Psi$ to facilitate a straightforward specification of coordination approaches for MDO frameworks.

    Lamb and Martins included objectives and constraint functions as nodes
    in an Extended DSM (XDSM) in order to capture a more complete description
    of solution strategies for MDAO problems \cite{Lambe2012}. Unlike REMS,
    XDSM also includes nodes for solvers and optimizers to enable complete
    definition of MDAO architectures. Martins uses XDSM to describe thirteen different
    optimization architectures in a survey paper that provides a novel
    classification of the different techniques \cite{Lambe2011}. With the
    additional information included in an XDSM, Lu and Martins applied both
    ordering and partitioning algorithms to an MDAO test problem termed the
    Scalable Problem \cite{Lu2012}.

### 2013-09-27 problem_formulation.tex
Submitted version of revisions 1

\begin{abstract}
The formulation of multidisciplinary design, analysis, and optimization (MDAO) problems has become increasingly complex as the number of analysis tools and design variables included in typical studies has grown.  This growth in the scale and scope of MDAO problems has been motivated by the need to incorporate additional disciplines and to expand the parametric design space to enable the exploration of unconventional design concepts.  In this context, given a large set of disciplinary analysis tools, the problem of determining a feasible data flow between tools to produce a specified set of system-level outputs is combinatorially challenging.   The difficulty is compounded in multi-fidelity problems, which are of increasing interest to the MDAO community.  In this paper, we propose an approach for addressing this problem based on the formalism of graph theory.  The approach begins by constructing the maximal connectivity graph (MCG) describing all possible interconnections between a set of analysis tools. Graph operations are then conducted to reduce the MCG to a fundamental problem graph (FPG) that describes the connectivity of analysis tools needed to solve a specified system-level design problem. The FPG does not predispose a particular solution procedure; any relevant MDO solution architecture could be selected to implement the optimization.  
The solution architecture can be represented in a problem solution graph (PSG).
The approach is applied to an example problem to formulate an FPG for a commercial aircraft MDAO study.
\end{abstract}

>>>>

\begin{abstract}
The formulation of multidisciplinary design, analysis, and optimization (MDAO) problems has become increasingly complex as the number of analysis tools and design variables included in typical studies has grown.  This growth in the scale and scope of MDAO problems has been motivated by the need to incorporate additional disciplines and to expand the parametric design space to enable the exploration of unconventional design concepts.  In this context, given a large set of disciplinary analysis tools, the problem of determining a feasible data flow between tools to produce a specified set of system-level outputs is combinatorially challenging.   The difficulty is compounded in multi-fidelity problems, which are of increasing interest to the MDAO community.  In this paper, we propose an approach for addressing this problem based on the formalism of graph theory.  The approach begins by constructing the maximal connectivity graph (MCG) describing all possible interconnections between a set of analysis tools. Graph operations are then conducted to reduce the MCG to a fundamental problem graph (FPG) that describes the connectivity of analysis tools needed to solve a specified system-level design problem. The FPG does not predispose a particular solution procedure; any relevant MDO solution architecture could be selected to implement the optimization.
Finally, the solution architecture can be represented in a problem solution graph (PSG). The graph approach is applied to an example problem based on a commercial aircraft MDAO study.
\end{abstract}
